### 2024-03-18 main.tex
Update on Overleaf.

%%
%% This is file `sample-acmsmall.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `all,journal,bibtex,acmsmall')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[acmsmall]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2024}
\acmYear{2024}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% These commands are for a JOURNAL article.
\acmJournal{JACM}
\acmVolume{37}
\acmNumber{4}
\acmArticle{111}
\acmMonth{8}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
%\usepackage{amssymb}
\usepackage{array}
\usepackage{minted}

% Configuration for minted package
\setminted{
    style=colorful, % Set code style to colorful
    fontsize=\scriptsize,
    linenos, % Enable line numbers
    breaklines, % Enable line breaks
    breakanywhere, % Allow line breaks anywhere
    autogobble,
}
%\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\newcommand\teo[1]{\textcolor{red}{teo:#1}}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Finch: A Datastructure-Driven Array Programming Language}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Willow Ahrens}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge}
  \state{Massachusetts}
  \country{USA}}
\email{willow@csail.mit.edu}

\author{Teodoro Fields Collin}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge}
  \state{Massachusetts}
  \country{USA}}
\email{teoc@mit.edu}

\author{Radha Patel}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge}
  \state{Massachusetts}
  \country{USA}}
\email{rrpatel@mit.edu}

\author{Kyle Deeds}
\affiliation{%
  \institution{University of Washington}
  \city{Seattle}
  \state{Washington}
  \country{USA}}
\email{kdeeds@cs.washington.edu}

\author{Changwan Hong}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge}
  \state{Massachusetts}
  \country{USA}}
\email{changwan@mit.edu}

\author{Saman Amarasinghe}
\affiliation{%
  \institution{MIT CSAIL}
  \city{Cambridge}
  \state{Massachusetts}
  \country{USA}}
\email{saman@csail.mit.edu}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Ahrens et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
From FORTRAN to Numpy, arrays have revolutionized how we express computation.  Arrays are the highest-performing datastructure with a long history of investment and innovation, from hardware support to compiler technology.  However, arrays can only handle dense rectilinear integer grids. Real world arrays often contain underlying structure, such as sparsity, runs of repeated values, or symmetry. We describe a compiler, Finch, which adapts existing programs and interfaces to the structure and sparsity of the inputs. Finch enables programmers to capture complex, real-world data scenarios with the same productivity they expect from dense arrays. Our approach enables new loop optimizations across multiple domains, unifying techniques such as sparse tensors, databases, and lossless compression. 
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>00000000.0000000.0000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
  Your, Paper}

\received{20 February 2007}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}

%- Fortran supported lots of control structures, no data structures, just dense arrays.
%
%- We tried to emulate complex data using only dense arrays but with complicated control structures.
%
%- Recently, people tried to build frameworks to support structured data, but gave up a lot of program side
%
%- We’re bringing these together, because if both of these can work together, we can more efficiently search the space of programs

\textbf{Finch makes high-performance computing easier with compilers that adapt programs to the structure of data}.  Currently, experts must hand write variations of even the simplest kernels, like matrix multiply, for each data structure and architecture to get performance. This makes performance engineering an expensive, repetitive task. Several popular programs compute the same thing on different data.

\textbf{Unstructured data is the easiest way to program, but the world is full of structured data}. Multidimensional dense (unstructured) arrays are the most important abstraction in computer science. Since their introduction in the 1950s with Fortran and Lisp, the simplicity of dense arrays has enabled extensive innovation, from compiler optimizations to convenience layers. However, structure is everywhere. Sparse arrays (which store only nonzero elements) can describe networks and simulations. Run-length encoding describes images, geometry and databases (such as a list of transactions with the date field all the same). Symmetry, bands, padding, and blocks are all natural properties which arise in the course of scientific computing, simulating the world around us.

\textbf{Structured data support is fragmented and incomplete}. Programmers today must choose between program flexibility and datastructure flexibility. Hand-written solutions are collected in diverse libraries like MKL, OpenCV, or SciPy. However, libraries will only ever support a subset of programs on a subset of datastructure combinations. Special-purpose compilers like TACO or CoRa support a select subset of data structures (only sparse, or only ragged arrays) in highly constrained classes of programs such as tensor contractions \cite{kjolstad_tensor_2019, fegade_cora_2022}.

\textbf{Finch handles the two main challenges for structured array data}:
\begin{itemize}
\item \textbf{Optimizations are specific to the indirection and patterns in datastructures}: These structures break the simple mapping between array elements and where they are stored in memory. For example, sparse arrays store lists of which coordinates are nonzero, whereas run-length-encoded arrays map several pixels to the same color value. These zero regions or repeated regions are optimization opportunities. Here, we show how Finch combines programs with datastructure descriptions to generate efficient, datastructure-specific code.

\item \textbf{Performance is highly structure dependent: Structure aware kernels are dramatically faster than unstructured ones, but the landscape of implementation decisions is also dramatically unpredictable}. For example, sparse kernels don’t need to compute on zeros, but this means that the precise input nonzero patterns act as computational filters, affecting the runtime as they interact with each other and the implementation. Finch allows you to search it more efficiently.
\end{itemize}

\subsection{Finch: Compiling Structured Data}
The finch array programming language adapts complex programs to array datastructures to produce efficient code. Array programming is the core abstraction behind many of the modern miracles of computing (e.g. neural networks, scientific simulation, database processing). Finch supports \textit{both} flexible programming constructs and diverse data structures. Finch supports loops, statements, if conditions, breaks, etc, over a wide variety of array structures, such as sparsity, run-length-encoding, symmetry, triangles, padding, or blocks.

Crucially, Finch is able to adapt simple, unstructured programs to the structure of the input data. It does this using Looplets, a new language that decomposes datastructure iterators hierarchically. Looplets represent the control-flow structures needed to iterate over them simultaneously with other structures in the program. Because looplets are compiled with progressive lowering, structure-specific mathematical optimizations such as integrals, multiply by zero, etc. can be implemented using simple compiler passes like term rewriting and constant propagation.

\subsection{Contributions}

\begin{enumerate}
\item More complex array structures than ever before. A complete level-by-level
structure-description language for expressing the structure of data
hierarchically. The first such set of formats to efficiently capture banded,
triangular, run-length-encoded, or sparse datasets, and any combination thereof.
\item A rich structured array programming language with for-loops
and complex control flow constructs at the same level of productivity
of dense arrays. To our knowledge, the Finch programming language is the first
to support if-conditions, early breaks, and multiple left hand sides over
structured data, as well as complex accesses such as affine indexing or scatter/gather.
\item A compiler that specializes programs to data structures 
predictably, facilitating an expressive language that makes it easier to search the complex
space of algorithms and datastructures. It is easy to implement new datastructures with a simple interface; the compiler analyzes the tensor lifecycle and inserts insert simple, overloadable, interface functions such as
initialization or finalization.
\item We evaluate the productivity of our language in several case studies,
showing that Finch can be used to accelerate a wide range of applications, 
from classic operations such as spmv and spgemm, to more complex applications such as image processing, graph analytics, and a high-level tensor operator fusion interface.
%We also demonstrate how Finch can fuse high-level operations to achieve a significant speedup over non-fused kernels. Additionally, as a case study, a high-level array programming language and fusion interface for operations such as map, broadcast, or reduce that can be compiled to efficient code using the previous loop-level abstractions.
%\item A complete set of level formats for expressing data patterns hierarchically in FiberTree-style decompositions. The first such set of formats to efficiently capture banded, triangular, run-length-encoded, or sparse-run-length-encoded datasets. The formats capture many use cases, from random updates to sequential construction.
%\item The Finch array language, mirroring simple for-loops with imperative code blocks and if-conditions. The first array programming language for the above data formats to support multiple outputs, affine indexing, and imperfectly-nested loops.
%\item Tensor lifecycles, a simple constraint on tensor reads and writes that elegantly restricts Finch programs to avoid complex data dependencies, and enables tensor polymorphism by providing implementers with well-defined functions to overload.
%\item Wrapper Tensors which modify existing datastructures and recombine them to support new patterns, such as affine indexing, padding, transposition, and slicing.
%\item Wrapper Levels which modify existing datastructures and enabling complex features such as atomic updates or contiguous versus separate allocation.
%\item We define the first mappings from the existing pydata/sparse array api high-level operations to low level finch notation
%\item <Performance Contributions>
\end{enumerate}

\begin{table}[h!]
\centering
\begin{tabular}{l|ccccc}
\textbf{Feature / Tool} & \textbf{Halide} & \textbf{Taco} & \textbf{Cora} & \textbf{Taichi} & \textbf{Finch} \\
\hline
Einsums and Contractions & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
Parallelism             & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
Multiple LHS            & \checkmark &            & \checkmark & \checkmark & \checkmark \\
Affine Indices          & \checkmark &            &            & \checkmark & \checkmark \\
Recurrence              & \checkmark &            &            &            &            \\
If-Conditions and Masks & \checkmark & \checkmark &            & \checkmark & \checkmark \\
Scatter Gather          & \checkmark &            &            & \checkmark & \checkmark \\
Early Break             &            & \checkmark &            &   \checkmark         & \checkmark \\
\end{tabular}
\caption{Feature support across various tools.}
\label{tab:features}
\end{table}

\begin{table}[h!]
\centering
\begin{tabular}{l|ccccc}
\textbf{Feature / Tool} & \textbf{Halide} & \textbf{Taco} & \textbf{Cora} & \textbf{Taichi} & \textbf{Finch} \\
\hline
Dense                    & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
Padded                   & \checkmark &            &            &            & \checkmark \\
One Sparse               &            & \checkmark &            & \checkmark & \checkmark \\
Sparse                   &            & \checkmark &            &            & \checkmark \\
Run-length               &            &            &            &            & \checkmark \\
Symmetric                &            &            &            &            & \checkmark \\
Regular Sparse Blocks    &            & \checkmark &            &            & \checkmark \\
Irregular Sparse Blocks  &            &            &            &            & \checkmark \\
Ragged                   &            &            & \checkmark &            & \checkmark \\
\end{tabular}
\caption{Support for various data structures across tools.}
\label{tab:data_structures}
\end{table}

\section{Background}
\subsection{Looplets}
\subsection{FiberTrees}

\subsection{Concordant Iteration}

\subsection{Protocols}

\section{The Finch Language}

\subsection{Syntax and Semantics}

\section{Bridging Looplets and Finch: The Tensor Interface}
\teo{I am doing stuff in this section.}
\input{sections/tensors.tex}

\subsection{Wrapper Tensors}

\subsection{Scalars}

\subsubsection{Sparse Scalars}
\subsubsection{Early Break Scalars}

\section{The Finch Compiler}

\subsection{Dimensionalization}

\subsection{Concordization}

\subsection{Bounds Analysis}

\subsection{Performance Warnings}

\subsection{Wrapperization}

\subsection{Simplification and Algebraic Transformations}

\section{Evaluation}

\subsection{Data-Driven Performance Engineering}
\subsubsection{Sparse-Sparse Matrix Multiply}

\begin{figure}
	\includegraphics[width=\linewidth]{erode4_speedup_over_opencv.png}
    \caption{Performance of Finch on erosion task (4 iterations).}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth]{spgemm_joel_speedup.png}
    \caption{A comparison of several matrix multiplication algorithms between Finch and Taco smaller matrices, ordered from small to big dimension. Note that inner products necessarily requires $O(n^2)$ work and taco's outer products format is dense. Finch can use a sparse outer products format and thus has an asymptotic advantage that becomes evident as the output dimensions grow.}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth]{spgemm_joel_speedup.png}
    \caption{A comparison of gustavson's algorithm between Finch and Taco on some larger matrices}
\end{figure}

Examples that demonstrate performance engineering in a datastructure-driven model

\subsubsection{SpMV}
Finch provides many flexible level formats to efficiently capture a variety of patterns in datasets, encompassing both the structure and type of data. More specifically, these formats can represent banded, triangular, run-length-encoded, blocked, hashed, and boolean data, as well as several other formats. Finch also provides the control flow necessary to manipulate the order in which data in these flexible level formats is read and written, enabling us to take advantage of multiple structural patterns concurrently—for example, we can exploit both sparsity and symmetry by using a sparse level format and restricting data reads to one triangle of a matrix. % I'm not sure if the right place for this is here - maybe should come earlier?

Structure in data arises both naturally, due to the chemical and physical properties of matter, and artificially via mathematical operations that induce particular patterns. Closely tailoring a storage format to a particular data pattern enables us to reduce the amount of stored values, make data accesses more efficient, and take advantage of spatial locality, resulting in more performant code.

The sparse matrix-vector multiplication kernel is a common operation in sparse linear algebra with many applications including conjugate gradients, graph algorithms, numerical analysis, and neural networks. The wide range of applications unsurprisingly results in a wide range of types of datasets making it an effective kernel to demonstrate the utility of having flexible data formats. In this case study, we highlight a few different Finch formats and the performance effects of conforming a dataset’s structure with its storage format, which Finch's datastructure-driven model enables us to do.

We compare Finch’s performance to that of TACO, SuiteSparseGraphBLAS, and Julia’s standard library.  We test using sparse matrices from a large selection of datasets spanning several previous papers: the datasets used by Vuduc et al. to test the OSKI interface [cite], Ahrens et al. to test a variable block row format partitioning strategy [cite], and Kjolstad et al. to test the TACO library [cite]. In addition, we also created several synthetic matrices containing bands or blocks of varying sizes as well as a permutation matrix to encapsulate a few additional use cases. The dense vector is randomly generated. We depict the performance of SpMV across the aforementioned tools and compare to the fastest Finch format for that particular dataset in Figures 1 and 2.

% TODO: make this be 3 columns
\begin{figure}
    \begin{minipage}[t]{0.315\textwidth}
        \vspace{0pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            @finch begin
                y .= 0
                for j = _, i = _
                    y[i] += A[i, j] * x[j]
                end
                return y
            end
        \end{minted}
    \end{minipage}%
    \begin{minipage}[t]{0.315\textwidth}
        \vspace{0pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            @finch begin
                y .= 0
                for j = _, i = _
                    y[j] += A[i, j] * x[i]
                end
                return y
            end
        \end{minted}
    \end{minipage}
    \begin{minipage}[t]{0.36\textwidth}
        \vspace{0pt} % Add this to ensure top alignment within minipage
        \begin{minted}{julia}
            @finch begin
                y .= 0
                for j = _
                    let x_j = x[j]
                        y_j .= 0
                        for i = _
                            let A_ij = A[i, j]
                                y[i] += x_j * A_ij
                                y_j[] += A_ij * x[i]
                            end
                        end
                        y[j] += y_j[] + diag[j] * x_j
                    end
                end
                return y
            end
        \end{minted}
    \end{minipage}
    \caption{Performance of SpMV across various tools.}
\end{figure}

\subsubsection{Tensor Formats}
We benchmarked the SpMV kernel using the Dense level format for the outer level and the SparseList, SparseVBL, SparseBand, and SparsePoint Finch levels for the inner level of the sparse matrix. To represent scalar values, we benchmarked with both the Element and Pattern levels. The SparseList, SparseVBL, and SparseBand formats were functional for all matrices, but the SparsePoint format was applicable only for the synthetic permutation matrices. The Element format was applicable to represent scalars in all matrices, while the Pattern level was applicable only for boolean matrices.

We found that the SpMV performance was superior for the level format that best paralleled the structure of the tensor. Namely, matrices with a clear blocked structure like $exdata\_1$, $TSOPF\_RS\_b678\_c1$, and $heart3$ performed notably well with the SparseVBL format in comparison to both TACO and the Finch SparseList formats (which we consider the closest analog to the sparse matrix format used in TACO and other libraries) with speedups of 2.16, X, and Y relative to TACO and of 1.47, X, and Y relative to the SparseList Finch SpMV implementation. Furthermore, the synthetic Toeplitz banded matrices we constructed performed the best with the SparseBand matrix, in particular with the $toeplitz\_large\_band$ and the $toeplitz\_medium\_band$ matrices having a speedup of A and B relative to TACO and a speedup of A and B relative to the SparseList Finch SpMV implementation.

There were also significant advantages of using the Pattern format instead of the Element format to represent scalar values in the matrices when these values were boolean. The matrices in the SNAP collection which represent graph datasets are boolean and had an average speedup of X over TACO and X over the same Finch program, but with an Element format.

\subsubsection{Symmetric SpMV}
Finch enables us to exploit symmetry in the sparse matrix of the SpMV kernel by providing the capabilities to reuse memory reads and insert control flow logic to restrict iterations to either the lower or upper triangle of the sparse matrix. We can apply this strategy with any level format. Every symmetric matrix in the SparseList and SparseList-Pattern formats has better performance when we use a Finch SpMV program that takes advantage of this symmetry. However, the regular Finch SpMV program has better performance for symmetric matrices than the symmetric Finch SpMV program for the other more specialized formats, likely because we need in-order accesses to fully capitalize on the specialized storage. Symmetric SpMV with the SparseList level format in Finch results in an average of 1.3x speedup over TACO and symmetric SpMV with the SparseList-Pattern format in Finch results in an average speedup of 1.15x over TACO . Notably, there is a 1.9x speedup for the HB/saylr4 matrix. 
\subsubsection{4D Blocked SpMV}

%Here's a figure with spmv_performance_sorted_(faster_than_taco).png and spmv_performance_sorted_(slower_than_taco).png

\begin{figure}
    \begin{minipage}[t]{0.5\textwidth}
        \vspace{0pt} % Add this to ensure top alignment within minipage
        \includegraphics[width=\linewidth]{spmv_performance_sorted_(faster_than_taco).png}
    \end{minipage}%
    \begin{minipage}[t]{0.5\textwidth}
        \vspace{0pt} % Add this to ensure top alignment within minipage
        \includegraphics[width=\linewidth]{spmv_performance_sorted_(slower_than_taco).png}
    \end{minipage}
    \caption{Performance of SpMV across various tools.}
\end{figure}

\begin{figure}
    \includegraphics[width=\linewidth]{spmv_performance_grouped.png}
    \caption{Performance of SpMV by Finch format.}
\end{figure}

\subsection{Programming over flexible data}

\subsubsection{Image Morphology}

\begin{figure}
	\includegraphics[width=\linewidth]{erode2_speedup_over_opencv.png}
    \caption{Performance of Finch on erosion task (2 iterations).}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth]{erode4_speedup_over_opencv.png}
    \caption{Performance of Finch on erosion task (4 iterations).}
\end{figure}

\begin{figure}
	\includegraphics[width=\linewidth]{hist_speedup_over_opencv.png}
    \caption{Performance of Finch on masked histogram task.}
\end{figure}

\subsubsection{Graph Analytics}
\begin{figure}
	\includegraphics[width=\linewidth]{bfs_speedup_over_graphs.jl.png}
	\includegraphics[width=\linewidth]{bellmanford_speedup_over_graphs.jl.png}
    \caption{Performance of graph apps across various tools.}
\end{figure}

Push-Pull BFS in Finch:
\begin{minted}{julia}
V = Tensor(Dense(Element(false)))
P = Tensor(Dense(Element(0)))
F = Tensor(SparseByteMap(Pattern()))
_F = Tensor(SparseByteMap(Pattern()))
A = Tensor(Dense(SparseList(Pattern())))
AT = Tensor(Dense(SparseList(Pattern())))

function finch_bfs_push_kernel(_F, F, A, V, P)
    @finch begin
        _F .= false
        for j=_, k=_
            if F[j] && A[k, j] && !(V[k])
                _F[k] |= true
                P[k] <<choose(0)>>= j #Only set the parent for this vertex
            end
        end
        return _F
    end
end

function finch_bfs_pull_kernel(_F, F, AT, V, P)
    p = ShortCircuitScalar{0}()
    @finch begin
        _F .= false
        for k=_
            if !V[k]
                p .= 0
                for j=_
                    if F[follow(j)] && AT[j, k]
                        p[] <<choose(0)>>= j #Only set the parent for this vertex
                    end
                end
                if p[] != 0
                    _F[k] |= true
                    P[k] = p[]
                end
            end
        end
        return _F
    end
end)
\end{minted}

\subsection{Implementing Numpy Array API in Finch}
\subsubsection{The Finch High-Level API (Needs a Name)}

\subsubsection{Finch Logic}

\subsubsection{Finch Interpreter}

\subsubsection{Lowering}
\subsubsection{Heuristic Optimization}

Find an example where fusing the python interface gives a big speedup over non-fused kernels.

%matmul, mttkrp, repeated ttm, triangle counting, multiple pointwise,
%in-place.
%dot((v^t .* u), w)) vs. 
%(v^t .* dot(u, w))

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
    To Mateusz, Hameer, and Jaeyeon for their excellent programming contributions to the Finch codebase.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography.bib}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\end{document}
\endinput