### 2024-03-27 sections/lang.tex
Update on Overleaf.

\section{The Finch Language}

\subsection{Syntax}

The syntax of Finch is displayed in Figure \ref{fig:syntax}. The Finch syntax
mirrors most imperative languages with for-loops and control flow. Notable
inclusions to the language at the level of statements includes \mintinline{julia}{for},
\mintinline{julia}{let}, blocks of code with \mintinline{julia}{if}, and 
the lifecycle functions that let us declare, freeze, and thaw tensors.
%
At the level of expressions, we support a wide array of scalar operations through literals and calls to functions whose properties and definitions are defined externally.
%
The expressions can also interact with indices and extents.
% FIXME: ???
Finally, as detailed in the previous section, tensors are defined externally via an interface that supports the declare, freeze, thaw, and unfurl functions.
%
The first three are supported directly in the syntax whereas the four will be introduced through evaluation of \mintinline{julia}{for}-loops loops and accesses,
in the next section.
%

Our syntax is highly permissive: by allowin blocks of code with multiple statements, we support both multiple outputs and temporary tensors, seen most prominently in our implementation of gustavson's algorithm for sparse-sparse matrix multiply and breadth-first-search.
%
Our breadth-first seach also uses an
\mintinline{julia}{if} statement to avoid operating on vertices outside the
frontier. 
%
We have found that \mintinline{julia}{if} is particularly useful in
our model for masking to avoid work.
%
However, to compile these constructs efficient, we must restrict this syntax in several ways, as we have alluded to in the previous section via life cycles.
%
In the semantics section, we detail other specifics that allow us to specialize our syntax efficient code over structured data.

% Finch uses looplets to lower \mintinline{julia}{for}-loops, though a sparse
% program with complicated looplets should produce output tensors with the same
% semantic value as a dense program with ordinary for-loops.  We include the
% semantics of how looplets are lowered as it affects the structure of the results
% and the asymptotic complexity of the program. Some applications view the
% structure of tensors as semantically meaningful.

% The \mintinline{julia}{let} statement (marked by the \finchdefine AST node)
% allows us to define tensor variables and reuse their value in multiple
% subexpressions without passing the variable through a tensor. Although it is too
% complex in this work to propagate constants through tensors, we can propagate
% constants (including 0) through a define statement. This allows the compiler to
% build up complicated functions of known operators and infer annihilation
% properties through them.It also allows use to eliminate common subexpressions,
% or tensor accesses, by reusing a tensor value in multiple places (such as the
% symmetric spmv in Figure \ref{fib:spmv_programs}). As another example, in the
% following program we don't need to define sparsity and annihilator properties of
% $f(a + b) * (a - b)$, as we can derive this from the more basic operators, and
% the code only reads from $A$ and $B$ once.
% \begin{minted}{julia}
%     let a = A[i, j]
%         let b = B[i, j]
%             C[i, j] = (a + b) * (a - b)
%         end
%     end
% \end{minted}

% We allow the use of blocks of code to describe multiple operations. This allows
% us to support both multiple outputs and temporary tensors, seen most prominently
% in our implementation of gustavson's algorithm for sparse-sparse matrix multiply
% and breadth-first-search. Our breadth-first seach also uses an
% \mintinline{julia}{if} statement to avoid operating on vertices outside the
% frontier. We have found that \mintinline{julia}{if} is particularly useful in
% our model for masking to avoid work.

% Finally, we add some explicit functions to advance the lifecycle of a tensor.
% Note that the user would not typically write \mintinline{julia}{@freeze} or
% \mintinline{julia}{@thaw} explicitly as those can be inferred through a separate
% pass.

\begin{figure}
\scriptsize
\noindent % This ensures the minipages fill the page width
\begin{minipage}{0.4\linewidth}
\begin{align*}
    \finchliteral(val \in \mathbb{V}) &:= \text{\mintinline{julia}{val}} \\
    \finchvalue(ex \in \mathbb{S}, type \in \mathbb{T}) &:= \text{\mintinline{julia}{ex :: type}} \\
    \finchtensor(name \in \mathbb{S}) &:= \text{\mintinline{julia}{name}} \\
    \finchindex(name \in \mathbb{S}) &:= \text{\mintinline{julia}{name}} \\
    \finchvar(name \in \mathbb{S}) &:= \text{\mintinline{julia}{name}} \\
    \finchextent(a \in E, b \in E) &:= \text{\mintinline{julia}{a : b}} \\
    \finchcall(f \in E, args\ldots \in E) &:= \text{\mintinline{julia}{f(args...)}} \\
    \finchaccess(tns \in T, idxs\ldots \in E) &:= \text{\mintinline{julia}{tns[idxs...]}} \\
    \finchfreeze(tns \in T) &:= \text{\mintinline{julia}{@freeze(tns)}} \\
    \finchthaw(tns \in T) &:= \text{\mintinline{julia}{@thaw(tns)}} \\
    \finchmode(tns \in T) &:= \text{\mintinline{julia}{@mode(tns)}} \\
\end{align*}
\end{minipage}%
\begin{minipage}{0.5\linewidth}
\begin{align*}
    \mathbb{V} &:= \text{the set of all values.} \\
    \mathbb{S} &:= \text{the set of all symbols.} \\
    \mathbb{T} &:= \text{the set of all types.} \\
    I &:= \finchindex \\
    A &:= \finchaccess \\
    V &:= \finchvar \\
    T &:= \finchtensor \\
    E &:= \finchliteral \mid \finchvalue \mid \finchindex \\
        &\quad \mid \finchvar \mid \finchextent \mid \finchcall \mid \finchaccess \\
    S &:= \finchassign \mid \finchloop \mid \finchdefine \\
        &\quad \mid \finchsieve \mid \finchblock \\
\end{align*}
\end{minipage}%
\begin{align*}
    \finchdeclare(tns \in T, init \in E, dims\ldots \in E) &:= \text{\mintinline{julia}{tns .= init(dims...)}} \\
    \finchassign(lhs \in A, op \in E, rhs \in E) &:= \text{\mintinline{julia}{lhs <<op>>= rhs}} \\
    \finchloop(idx \in I, range \in E, body \in S) &:= \text{\mintinline{julia}{for idx = range; body end}} \\
    \finchdefine(var \in V, val \in E, body \in S) &:= \text{\mintinline{julia}{let var = val; body end}} \\
    \finchsieve(cond \in E, body \in S) &:= \text{\mintinline{julia}{if cond; body end}} \\
    \finchblock(bodies\ldots \in S) &:= \text{\mintinline{julia}{begin; bodies... end}}
\end{align*}
\caption{The syntax of the finch language. Note that instead of a classic EBNF
form, we have constrained the classes of subterms on the left hand side to allow
for descriptive names of the arguments. Compare this grammar to the Concrete
Index Notation of TACO \cite[Figure~3]{kjolstad_tensor_2019}, noting the
addition of multiple left hand sides through code blocks, and the explicit
initialization and finalization of tensors.}\label{fig:syntax}
\end{figure}

\subsection{Semantics}

We present a sketch of a small-step operational semantics, showing how to execute a Finch program in a host language.
%
We distinguish evaluation of the language by evolving compiler state with $\langle \rangle$ and evaluation in the host language via $\llangle\rrangle$.
%
For example, in Figure~\ref{fig:semantics_core}, we declare the core semantics of the language and the Define rule (\mintinline{julia}{let}) adds a definition to the compiler state whereas the Call rule passes evaluation off to the host.
%
In Figure~\ref{fig:semantics_looplets}, we detail the Looplet evaluation semantics, which details how acceses within a loop can be replaced with Loopets, which are progressively rewritten to host code.
%
Due to the use of rewrite rules, our semantics here make use of an evaluation context.

% We break our description of the semantics of Finch into three parts, those
% relating to lowering the language itself in Figure~\ref{fig:semantics_core},
% those pertaining to the lifecycle of tensors in
% Figure~\ref{fig:semantics_lifecycle}, and the semantics of the looplets in
% Figure~\ref{fig:semantics_looplets}. Note that in addition to the rules seen
% here, we also apply extensive symbolic simplification of the program whenever we
% can. In the implementation of the language we have heuristics for when to run a
% simplification pass, but for the purposes of understanding the semantics of the
% language we should understand simplification rules as occurring whenever possible
% before continuing to execute the program as normal.

% Note that both Looplets and Finch are designed to be embedded into a larger
% language, and a separate state variable is introduced for this language.

\begin{figure}
    \centering
    \footnotesize

    \begin{prooftree}
    \hypo{\langle val, (e, t) \rangle
    \rightarrow val'}
    \hypo{var\not\in e}
    \infer2[$Define$]{\splitfrac{\langle\finchdefine(var, val, body), (e, t)\rangle}{\rightarrow \langle body, (e[var \mapsto val'], t) \rangle}}
    \end{prooftree}
    \hfill
    \begin{prooftree}
        \hypo{\langle args_i, (e, t) \rangle \Rightarrow vals_i}
        \hypo{\langle f, (e, t) \rangle \Rightarrow g}
        \infer2[$Call$]{\langle\finchcall(f, args...), (e, t)\rangle \rightarrow \llangle g(vals...), t \rrangle}
    \end{prooftree}
    \vspace{6pt}

    \begin{prooftree}
        \hypo{}
        \infer1[$Literal$]{\langle\finchliteral(val), (e, t)\rangle \rightarrow val}
    \end{prooftree}
    \hfill
    \begin{prooftree}
        \hypo{}
        \infer1[$Value$]{\langle\finchvalue(ex, type), (e, t)\rangle \Rightarrow \llangle ex, t \rrangle}
    \end{prooftree}
    \vspace{6pt}

    \begin{prooftree}
        \infer0[$Variable$]{\splitfrac{\langle\finchvar(name), (e, t)\rangle}{\rightarrow e(\finchvar(name))}}
    \end{prooftree}
    \hfill
    \begin{prooftree}
        \infer0[$Index$]{\splitfrac{\langle\finchindex(name), (e, t)\rangle}{\rightarrow e(\finchindex(name))}}
    \end{prooftree}
    \vspace{6pt}

    \begin{prooftree}
        \hypo{\langle head, s_1 \rangle \rightarrow s_2}
        \infer1[$BlockHead$]{\langle\finchblock(head, tail...), s_1 \rangle \rightarrow \langle \finchblock(tail...), s_2 \rangle}
    \end{prooftree}
    \hfill
    \begin{prooftree}
        \infer0[$EmptyBlock$]{\langle\finchblock(), s\rangle \rightarrow s}
    \end{prooftree}
    \vspace{6pt}

    \begin{prooftree}
        \hypo{\langle cond, s \rangle \Rightarrow true}
        \infer1[$SieveTrue$]{\langle\finchsieve(cond, body), s\rangle \rightarrow \langle body, s \rangle}
    \end{prooftree}
    \hfill
    \begin{prooftree}
        \hypo{\langle cond, s \rangle \Rightarrow false}
        \infer1[$SieveFalse$]{\langle\finchsieve(cond, body), s\rangle \rightarrow s}
    \end{prooftree}
    \vspace{6pt}

    \begin{prooftree}  
    \hypo{s = (e, t)}
    \hypo{\finchmode(tns)\not\in e}
    \hypo{e(tns) = tns'}
    \hypo{\langle init, s \rangle \Rightarrow init'}
    \hypo{\forall i \langle init, dims_i \rangle \Rightarrow dims'_i}
    \infer5[$Declare$]{\langle \finchdeclare(tns, init, dims), (e, t)\rangle \rightarrow (e [\finchmode(tns) \mapsto update], \llangle declare(tns', init', dims'...), t \rrangle)}
    \end{prooftree}
    \vspace{6pt}

    \begin{prooftree}  
    \hypo{s = (e, t)}
    \hypo{e(\finchmode(tns)) = update}
    \hypo{e(tns) = tns'}
    \infer3[$Freeze$]{\langle\finchfreeze(tns), (e, t)\rangle \rightarrow (e [\finchmode(tns) \mapsto read], \llangle freeze(tns'), t \rrangle)}
    \end{prooftree}
    \vspace{6pt}

    \begin{prooftree}  
    \hypo{s = (e, t)}
    \hypo{e(\finchmode(tns)) = read}
    \hypo{e(tns) = tns'}
    \infer3[$Thaw$]{\langle\finchthaw(tns), s\rangle \rightarrow (e [\finchmode(tns) \mapsto update], \llangle thaw(tns'), t \rrangle)}
    \end{prooftree}

    \caption{Basic evaluation semantics, roughly defining most of these language
    constructs to function similarly to their classical definitions. Note that
    The state $s$ of the finch compiler is a tuple $(e, t)$ of a variable value
    environment and another state $t$ corresponding to the state in the
    host language. Several Looplets introduce variables into the embedding
    language, which may be read when evaluating the \finchvalue node.
    %
    The state $s$ of the finch compiler
    is a tuple $(e, t)$ of a variable value environment and another state $t$
    corresponding to the state in the host language.
    This means that rules which modify $t$ are running in the host language. All of the lifecycle
    functions are designed to be implemented and executed in the host language,
    but these semantics enforce that each of these functions may update state in
    the host language and flip the mode of the tensor between read and update}
    \label{fig:semantics_core}
\end{figure}

\begin{figure}
    \centering
    \footnotesize

\begin{multline*}
E := [\cdot] | \finchloop(idx, ext, E | body) | \finchblock(E | head, E | tail) | \finchsieve(E | cond, E | body) |\\ 
\finchdeclare(var, E|rhs, E|body) | \finchcall(E|f, E|args...) | \finchaccess(tns, E | idxs...)
\end{multline*}

\begin{prooftree}
    \hypo{e(tns) \mapsto tns'}
    \hypo{\llangle unfurl(tns', ext)\rrangle \mapsto tns''}
    \infer2[$Unfurl$]{\langle \finchloop(i, ext, E[\finchaccess(tns, j..., i)]), s\rangle \rightarrow \langle \finchloop(i, ext, E[\finchaccess(tns'', j..., i)]), s\rangle}
\end{prooftree}
\vspace{6pt}

\begin{prooftree}
    \infer0[$Run$]{\splitfrac{\langle \finchloop(i, ext, E[\finchaccess(run(body), j..., i)]), s\rangle}{\rightarrow \langle \finchloop(i, ext, E[\finchaccess(body, j...)]), s\rangle}}
\end{prooftree}
\hfill
\begin{prooftree}
    \hypo{e(i) = i'}
    \hypo{\llangle seek(i'), t \rrangle \rightarrow t'}
    \infer2[$Lookup$]{\splitfrac{\langle E[\finchaccess(lookup(seek, body), j..., i)], (e, t)\rangle}{\rightarrow \langle E[\finchaccess(body, j...)], (e, t')\rangle}}
\end{prooftree}
\vspace{6pt}

\begin{prooftree}
    \hypo{\llangle cond, t \rrangle \Rightarrow true}
    \infer1[]{\splitfrac{\langle E[\finchaccess(switch(cond, head, tail), i...)], s\rangle}{\rightarrow \langle E[\finchaccess(head, i...)], s\rangle}}
\end{prooftree}
\hfill
\begin{prooftree}
    \hypo{\llangle cond, t \rrangle \Rightarrow false}
    \infer1[$Switch$]{\splitfrac{\langle E[\finchaccess(switch(cond, head, tail), i...)], s\rangle}{\rightarrow \langle E[\finchaccess(tail, i...)], s\rangle}}
\end{prooftree}
\vspace{6pt}

\begin{prooftree}
    \infer0[$Phase$]{\splitfrac{\langle \finchloop(i, extent(a, b), E[\finchaccess(phase(extent(c, d), body), j..., i)]), s\rangle}{\rightarrow \langle \text{\finchloop}(i, extent(max(a, c), min(b, d)), E[\text{\finchaccess}(body, j..., i)]), s\rangle}}
\end{prooftree}
\vspace{6pt}

\begin{prooftree}
    \hypo{\llangle preamble, t \rrangle \rightarrow t'}
    \hypo{\langle E[body], (e, t') \rangle \rightarrow (e', t'')}
    \hypo{\llangle epilogue, t'' \rrangle \rightarrow t'''}
    \infer3[$Thunk$]{\langle E[thunk(preamble, body, epilogue)], (e, t')\rangle \rightarrow (e', t''')}
\end{prooftree}
\vspace{6pt}

\begin{prooftree}
    \hypo{\langle \finchloop(i, ext, E[\finchaccess(head, j..., i)]), s\rangle \rightarrow s'}
    \infer1[$Sequence$]{\splitfrac{\langle \finchloop(i, ext, E[\finchaccess(sequence(head, tail), j..., i)]), s\rangle}{\rightarrow \langle \finchloop(i, ext, E[\finchaccess(tail, j..., i)]), s'\rangle}}
\end{prooftree}
\hfill
\begin{prooftree}
    \hypo{\langle node, algebra \rangle \rightarrow node'}
    \infer1[$Simplify$]{\langle E[node], s\rangle \rightarrow \langle E[node'], s\rangle}
\end{prooftree}
\vspace{6pt}

\begin{prooftree}
    \hypo{\llangle seek(a), t \rrangle \rightarrow t'}
    \infer1[$StepperSeek$]{\splitfrac{\langle \finchloop(i, extent(a, b), E[\finchaccess(stepper(seek, body, next), j..., i)]), (e, t)\rangle}{\rightarrow \langle \finchloop(i, extent(a, b), E[\finchaccess(stepper(body, next), j..., i)]), (e, t')\rangle}}
\end{prooftree}
\vspace{6pt}

\begin{prooftree}
    \hypo{\langle \finchloop(i, ext, E[\finchaccess(body, j..., i)]), (e, t) \rangle \rightarrow (e', t')}
    \hypo{\llangle next, t' \rrangle \rightarrow t''}
    \infer2[$StepperNext$]{\splitfrac{\langle \finchloop(i, ext, E[\finchaccess(stepper(body, next), j..., i)]), s\rangle}{\rightarrow \langle \finchloop(i, ext, E[\finchaccess(stepper(body, next), j..., i)]), (e', t'')\rangle}}
\end{prooftree}

    \caption{Looplet evaluation semantics. 
    The state $s$ of the finch compiler is a tuple $(e, t)$ of a variable value
    environment and another state $t$ corresponding to the state in the host
    language. This means that rules which modify $t$ are running in the host language. Note that $E$ is an evaluation context that applies anywhere in
    the syntax tree. The nonlocal evaluations of Looplets are what allow looplets to
    hoist conditions and subranges out of loops. However, this also means we must specify
    the priority in which we apply looplet rules, which is as follows:
    $Thunk > Phase > Switch > Simplify > Run > Spike > Sequence > StepperSeek > StepperNext > Lookup > Unfurl$
    Many looplets, most notably the thunk looplet, introduce variables into the
    host language environment.  While variables introduced by a looplet may be
    modified by the looplet itself (steppers often increment some state
    variables), we forbid child looplets from modifying any state variables that
    they didn't introduce. This allows us to treat the \finchvalue node as a
    constant.
    %
    Note: the $Simplify$ rule references $algebra$, which is our variable
    defining a set of straightforward simplification rules. These rules include
    simple properties like $x * 0 \rightarrow 0$ to more complicated ones such
    as constant propagation. We omit the full set of rules for brevity, but
    point the curious reader to \cite[Figure 5]{ahrens_looplets_2023} for some examples}
    \label{fig:semantics_looplets}
\end{figure}