### 2024-04-02 sections/abstract.tex
updates

\begin{abstract}
From FORTRAN to Numpy, arrays have revolutionized how we express computation. However, arrays in these, and almost all prominent systems, can only handle dense rectilinear integer grids.
Real world arrays often contain underlying structure, such as sparsity, runs of repeated values, or symmetry.
Support for structured data is fragmented and incomplete.
Existing frameworks limit the array structures and program control flow they support to better simplify the problem.

In this work, we propose a new programming language, Finch, which supports \textit{both}
flexible control flow and diverse data structures. Finch facilitates a programming model which
resolves the challenges of computing over structured arrays by combining control flow and data
structures into a common representation where they can be co-optimized. In particular,
Finch automatically specializes the control flow to the data so that performance engineers can
focus on experimenting with many algorithms. Finch supports a familiar programming language
of loops, statements, if conditions, breaks, etc., over a wide variety of array structures, such as
sparsity, run-length-encoding, symmetry, triangles, padding, or blocks. Finch reliably utilizes the key
properties of structure, such as structural zeros, repeated values, or clustered non-zeros. We show that this leads to dramatic speedups in operations such as spmv and spgemm, image processing, graph analytics, and a high-level tensor operator fusion interface.
\end{abstract}