### 2024-04-04 sections/background.tex
Update on Overleaf.

\section{Background On Looplets}
Finch represents iteration patterns using Looplets, a language that decomposes datastructure iterators hierarchically. 
%
Looplets represent the control-flow structures needed to iterate over any given datastructure, or multiple datastructures simultaneously. 
%
In particular, looplets are good at lifting code to the highest loop level that it's needed and subdividing iteration hierarchically in coordinate space.
%
Because looplets are compiled with progressive lowering, structure-specific mathematical optimizations such as integrals, multiply by zero, etc. can be implemented using simple compiler passes like term rewriting and constant propagation during the intermediate lowering stages \cite{ahrens_looplets_2023}.

The looplets are described in Figure~\ref{fig:looplets}. We simplify the presentation to focus on the semantics, rather than precise implementation.  Several looplets introduce or modify variables in the scope of the target language. It is assumed that if a looplet introduces a variable to be used in a child looplet, the child looplet will not modify that variable.

% \begin{figure}[ht]
%     \footnotesize
%     \begin{minipage}[c]{0.65\linewidth}
%         $\finchlookup(seek, body)$: The Lookup Looplet represents a
%         randomly accessible region of an iterator. The body of the lookup is
%         understood to have one less dimension than the lookup itself, as we have
%         already ``looked up'' that index in the tensor by the time we reach the
%         body. \texttt{seek(i)} is a function that updates state to the given
%         index.
%     \end{minipage}%
%     \begin{minipage}[c]{0.35\linewidth}
%         \centering
%         \includegraphics[scale=0.20]{Looplets-lookup.png}
%     \end{minipage}
%     \vspace{3pt}

%     \begin{minipage}[c]{0.65\linewidth}
%         $\finchrun(body)$: The Run Looplet represents a constant
%         region of an iterator. The body of the run is understood to have one
%         less dimension than the lookup itself, as all of the bodies are
%         identical.
%     \end{minipage}%
%     \begin{minipage}[c]{0.35\linewidth}
%         \centering
%         \includegraphics[scale=0.20]{Looplets-run.png}
%     \end{minipage}
%     \vspace{3pt}

%     \begin{minipage}[c]{0.65\linewidth}
%         $\finchphase(c:d, body)$: The Phase Looplet represents a
%         restriction of the range on which a loop should execute, and allows us
%         to succinctly express the ranges on which children of compound looplets
%         are defined.
%     \end{minipage}%
%     \begin{minipage}[c]{0.35\linewidth}
%         \centering
%         \includegraphics[scale=0.20]{Looplets-phase.png}
%     \end{minipage}
%     \vspace{3pt}

%     \begin{minipage}[c]{0.65\linewidth}
%         $\finchswitch(cond, head, tail)$: The Switch Looplet allows
%         us to specialize the body of a looplet based on a condition, evaluated
%         in the embedding context. If the condition is true, we use `head`,
%         otherwise we use `tail`. Switch has a high lowering priority so we can
%         see what's inside of it and lower that appropriately. This also lifts
%         the condition as high as possible into the loop nest. The condition is
%         assumed to evaluate to a boolean.
%     \end{minipage}%
%     \begin{minipage}[c]{0.35\linewidth}
%         \centering
%         \includegraphics[scale=0.20]{Looplets-switch.png}
%     \end{minipage}
%     \vspace{3pt}

%     \begin{minipage}[c]{0.65\linewidth}
%         $\finchthunk(preamble, body, epilogue)$: The Thunk Looplet
%         allows us to cache certain computations in the state under which the
%         body will execute. This is useful for computing and caching the results
%         of expensive computations.
%     \end{minipage}%
%     \begin{minipage}[c]{0.35\linewidth}
%         \centering
%         \includegraphics[scale=0.20]{Looplets-thunk.png}
%     \end{minipage}
%     \vspace{3pt}

%     \begin{minipage}[c]{0.65\linewidth}
%         $\finchsequence(head, tail)$: The Sequence looplet represents the
%         concatenation of two looplets. Both arguments must be phase looplets, and
%         are assumed to be nonoverlapping, covering, and in order.
%     \end{minipage}%
%     \begin{minipage}[c]{0.35\linewidth}
%         \centering
%         \includegraphics[scale=0.20]{Looplets-sequence.png}
%     \end{minipage}
%     \vspace{3pt}

%     \begin{minipage}[c]{0.65\linewidth}
%         $\finchspike(body, tail)$ The Spike Looplet represents a run
%         followed by a single value. In this paper, Spike will be considered a
%         shorthand for $\finchsequence(\finchphase(i:j-1, \finchrun(body)),
%         \finchphase(j:j, \finchrun(tail)))$.  In the Finch compiler, spikes are
%         handled with special care, since they are an opportunity to align the
%         final run to the end of the root loop extent, without using any special
%         bounds inference.
%     \end{minipage}%
%     \begin{minipage}[c]{0.35\linewidth}
%         \centering
%         \includegraphics[scale=0.20]{Looplets-spike.png}
%     \end{minipage}
%     \vspace{3pt}

%     \begin{minipage}[c]{0.65\linewidth}
%         $\finchstepper([seek], next, body)$ The stepper looplet
%         represents a variable number of looplets, concatenated. Since our
%         looplets may be skipped over due to conditions or various rewrites, the
%         $seek$ function allows us to fast-forward the state to the start of the
%         root loop extent when it comes time to lower the stepper. The $next$
%         function advances the state to the next iteration of the stepper. 
%     \end{minipage}%
%     \begin{minipage}[c]{0.35\linewidth}
%         \centering
%         \includegraphics[scale=0.20]{Looplets-stepper.png}
%     \end{minipage}
%     \caption{The looplet language, as understood in a correct execution of a Finch program.}
%     \vspace{-8pt}
% \end{figure}

\begin{figure}[ht]
\begingroup
\footnotesize
\setlength{\tabcolsep}{10pt} 
\renewcommand{\arraystretch}{1.5}
\begin{tabular} {|l|c|} 
\hline
    \begin{minipage}[c]{0.65\linewidth}
        $\finchlookup(seek, body)$: The Lookup Looplet represents a
        randomly accessible region of an iterator. The body of the lookup is
        understood to have one less dimension than the lookup itself, as we have
        already ``looked up'' that index in the tensor by the time we reach the
        body. \texttt{seek(i)} is a function that updates state to the given
        index.
    \end{minipage} &
    \begin{minipage}[c]{0.25\linewidth}
        \centering
        \includegraphics[scale=0.20]{Looplets-lookup.png}
    \end{minipage} \\ \hline
    \begin{minipage}[c]{0.65\linewidth}
        $\finchrun(body)$: The Run Looplet represents a constant
        region of an iterator. The body of the run is understood to have one
        less dimension than the lookup itself, as all of the bodies are
        identical.
    \end{minipage} &
    \begin{minipage}[c]{0.25\linewidth}
        \centering
        \includegraphics[scale=0.20]{Looplets-run.png}
    \end{minipage} \\ \hline
    \begin{minipage}[c]{0.65\linewidth}
        $\finchphase(c:d, body)$: The Phase Looplet represents a
        restriction of the range on which a loop should execute, and allows us
        to succinctly express the ranges on which children of compound looplets
        are defined.
    \end{minipage} &
    \begin{minipage}[c]{0.25\linewidth}
        \centering
        \includegraphics[scale=0.20]{Looplets-phase.png}
    \end{minipage} \\ \hline
    \begin{minipage}[c]{0.65\linewidth}
        $\finchswitch(cond, head, tail)$: The Switch Looplet allows
        us to specialize the body of a looplet based on a condition, evaluated
        in the embedding context. If the condition is true, we use `head`,
        otherwise we use `tail`. Switch has a high lowering priority so we can
        see what's inside of it and lower that appropriately. This also lifts
        the condition as high as possible into the loop nest. The condition is
        assumed to evaluate to a boolean.
    \end{minipage} &
    \begin{minipage}[c]{0.25\linewidth}
        \centering
        \includegraphics[scale=0.20]{Looplets-switch.png}
    \end{minipage} \\ \hline
    \begin{minipage}[c]{0.65\linewidth}
        $\finchthunk(preamble, body, epilogue)$: The Thunk Looplet
        allows us to cache certain computations in the state under which the
        body will execute. This is useful for computing and caching the results
        of expensive computations.
    \end{minipage} &
    \begin{minipage}[c]{0.25\linewidth}
        \centering
        \includegraphics[scale=0.20]{Looplets-thunk.png}
    \end{minipage} \\ \hline
    \begin{minipage}[c]{0.65\linewidth}
        $\finchsequence(head, tail)$: The Sequence looplet represents the
        concatenation of two looplets. Both arguments must be phase looplets, and
        are assumed to be nonoverlapping, covering, and in order.
    \end{minipage} &
    \begin{minipage}[c]{0.25\linewidth}
        \centering
        \includegraphics[scale=0.20]{Looplets-sequence.png}
    \end{minipage} \\ \hline
    \begin{minipage}[c]{0.65\linewidth}
        $\finchspike(body, tail)$ The Spike Looplet represents a run
        followed by a single value. In this paper, Spike will be considered a
        shorthand for $\finchsequence(\finchphase(i:j-1, \finchrun(body)),
        \finchphase(j:j, \finchrun(tail)))$.  In the Finch compiler, spikes are
        handled with special care, since they are an opportunity to align the
        final run to the end of the root loop extent, without using any special
        bounds inference.
    \end{minipage} &
    \begin{minipage}[c]{0.25\linewidth}
        \centering
        \includegraphics[scale=0.20]{Looplets-spike.png}
    \end{minipage} \\ \hline
    \begin{minipage}[c]{0.65\linewidth}
        $\finchstepper([seek], next, body)$ The stepper looplet
        represents a variable number of looplets, concatenated. Since our
        looplets may be skipped over due to conditions or various rewrites, the
        $seek$ function allows us to fast-forward the state to the start of the
        root loop extent when it comes time to lower the stepper. The $next$
        function advances the state to the next iteration of the stepper. 
    \end{minipage} &
    \begin{minipage}[c]{0.25\linewidth}
        \centering
        \includegraphics[scale=0.20]{Looplets-stepper.png}
    \end{minipage} \\\hline
\end{tabular}
\endgroup
\vspace{-8pt}
\caption{The looplet language, as understood in a correct execution of a Finch program.}
\end{figure}